---
title: "ABC_Bembix_simple"
author: "Femke Batsleer"
output: html_document
---

#ABC analysis
##Loading data
Data are loaded, created by other R-scripts. Field data and for each scenario: parameters and summary statistics.  
The data are adjusted to make sure they have the right format for the analysis (True and False become 0 and 1, response functions get numbers etc. because everything needs to be numeric).  

###Summary statistics  
All summary stats are loaded and a restricted dataset is created to include the most important ones (the ones that we want to select models on).  
```{r, message=FALSE, warning=FALSE}
setwd("D:/fbatslee/OneDrive - UGent/WP 2 Modelling Thesis/2 - Analysis/ABC4IBM/ABC")

library(abc)
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)
library(gridExtra)

#loading the data: summary stats, params scenarios, field data summary stats
##summary stats
getwd()
ss_random <- read.table("../Data outputs/Simple 20-11 INLA/Summary_stats_random_2019-11-21.txt", header=T)
ss_bimodal <- read.table("../Data outputs/Simple 20-11 INLA/Summary_stats_bimodal_2019-11-28.txt", header=T)
ss_BU <- read.table("../Data outputs/Simple 20-11 INLA/Summary_stats_bottom-up_2019-11-28.txt", header=T)
ss_CF <- read.table("../Data outputs/Simple 20-11 INLA/Summary_stats_CF_2020-04-23.txt", header=T)
ss_CA <- read.table("../Data outputs/Simple 20-11 INLA/Summary_stats_conspec-attraction_2019-11-28.txt", header=T)
ss_PEBU <- read.table("../Data outputs/Simple 20-11 INLA/Summary_stats_prev-exp-BU_2019-11-28.txt", header=T)

##combine the datasets
ss_all <- plyr::rbind.fill(ss_random, ss_bimodal, ss_CA, ss_CF, ss_BU, ss_PEBU)

colnames(ss_all) <- c("scenario", "file title", 'RK2', 'RK5', 'RK10', 'RK15', 'RK20', 'RK30', 'RK40',
                      'MC0', 'MC2', 'MC4', 'MC6', 'MC8', 'MC10', 'MC12', 'MC14', "NA_internal_loops",
                      "NA_all_loops","NA_dens_undirected", "NA_dens_directed",
                      "NA_reciproc","NA_transitivity_und","NA_transitivity_dir")

str(ss_all)
```


###Field data
Master thesis data  
```{r}
##field data summary stats
ss_field_df <- read.table("../Field Summary stats/Summary_stats_field.txt", header=T)
ss_field_df$X0 <- NULL
ss_field_df$scenario <- NULL
ss_field_df[] <- lapply(ss_field_df[], as.numeric)
colnames(ss_field_df) <- c("mean_RK", "sd_RK", 'RK2', 'RK5', 'RK10', 'RK15', 'RK20', 'RK30', 'RK40',
                           'mean_MC', 'sd_MC', 'MC0', 'MC2', 'MC4', 'MC6', 'MC8', 'MC10',
                           'MC12', 'MC14', 'MC16', 'MC18', 'MC20',
                           "NA_internal_loops", "NA_all_loops","NA_dens_undirected", "NA_dens_directed",
                           "NA_reciproc","NA_transitivity_und","NA_transitivity_dir")
ss_field_df$mean_MC <- NULL
ss_field_df$mean_RK <- NULL
ss_field_df$sd_MC <- NULL
ss_field_df$sd_RK <- NULL
ss_field_df$MC16 <- NULL
ss_field_df$MC18 <- NULL
ss_field_df$MC20 <- NULL
#make a 1D vector of it
ss_field <- as.numeric(ss_field_df[1,])

# #same for the restricted summary stats [c(1, 4:10, 13:23, 24:ncol(ss_all))]
# ss_field_res_df <- ss_field_df[c(1:3, 5:12, 19:20, 22:23, 25)]
# ss_field_res <- as.numeric(ss_field_res_df[1,])
# ss_field_res_df
```

###Parameters
Model parameters for all the runs.  
```{r, message=FALSE, warning=FALSE}
##priors/parameters of the models
#2 --> with weight day and max dens
params_random <- read.table("../Data outputs/Simple 20-11 INLA/Params_random_2019-11-21.txt", header=T)
params_bimodal <- read.table("../Data outputs/Simple 20-11 INLA/Params_bimodal_2019-11-28.txt", header=T)
params_BU <- read.table("../Data outputs/Simple 20-11 INLA/Params_bottom-up_2019-11-28.txt", header=T)
params_CF <- read.table("../Data outputs/Simple 20-11 INLA/Params_CF_2020-04-23.txt", header=T)
params_CA <- read.table("../Data outputs/Simple 20-11 INLA/Params_conspec-attraction_2019-11-28.txt", header=T)
params_PEBU <- read.table("../Data outputs/Simple 20-11 INLA/Params_prev-exp-BU_2019-11-28.txt", header=T)

##these dataframes are converted to numeric ones (True-->1, False-->0,...)
change_df_num <- function(df){
  df$scenario <- NULL
  df[] <- lapply(df[], as.character)
  df[df=='True'] <- '1'
  df[df=='False'] <- '0'
  df[df=='response_par'] <- '1'
  df[df=='response_type2'] <- '2'
  df[] <- lapply(df[], as.numeric)
  return(df)
}

params_bimodal <- change_df_num(params_bimodal)
params_BU <- change_df_num(params_BU)
params_CF <- change_df_num(params_CF)
params_CA <- change_df_num(params_CA)
params_PEBU <- change_df_num(params_PEBU)

##combine them in 1 dataset
params_all <- plyr::rbind.fill(params_random,params_bimodal, params_CA, params_CF, params_BU, params_PEBU)
head(params_all)
```

###Indexes
A vector has to be created with the names in of the models corresponding to the different runs.  

```{r}
#indexes -> which row corresponds with which model
indexes <- as.vector(unlist(ss_all[,1]))
str(indexes)

#keep summary stats with scenario attached to df
ss_all_w_scen <- ss_all
# ss_res_w_scen <- ss_res
#make sure the summary stats are numeric values (scenario will be replaced with 1..6)
ss_all[] <- lapply(ss_all[], as.numeric) #make sure all are numeric values
# ss_res[] <- lapply(ss_res[], as.numeric)

ss_all$scenario <- NULL #take scenario out of summary stats, to make calculations easier
ss_all$`file title` <- NULL
# ss_res$scenario <- NULL
```


####Correlation between summary stats
Within the Ripley's K, there is high correlation. But less for the high values. We should select distances that are further apart from each other. F.i. RK2, RK10, RK20, RK30, RK40.  
Within Mark correlation, there is no real correlation. But about 5 will be selected (to be equal to amount of RK and NM).  
Within Network metrics, there is some correlation. Especially for the same metric for undirected and directed networks. These will be dropped.  
For the correlation betwoon RKs and MCs, there doesn't seem to be a big correlation, but the restriction of certain values (low RK40 or high RK15-20-30, there are only values around 0 possible for MC0) can be seen.  
For correlation between MCs and network metrics, there is no correlation.  
For the network metrics and Ripley's K, there is some structure. But this is expected with such high level network metrics. Variation seems to be high enough.  

```{r}
#Within Ripley's K
pairs(ss_all[sample(nrow(ss_all),1000),c(1:7)])
#within Mark correlation
pairs(ss_all[sample(nrow(ss_all),1000),c(8:15)])
#within network metrics
pairs(ss_all[sample(nrow(ss_all),1000), c(16:ncol(ss_all))])

pairs(ss_all[sample(nrow(ss_all),1000),c(4:10)])
pairs(ss_all[sample(nrow(ss_all),1000),c(8:10, 16,17,19,20,22)])
pairs(ss_all[sample(nrow(ss_all),1000),c(1,3,5,6,7, 16,17,19,20,22)])

#make selection
ss_all_sel <- ss_all %>% select("RK2","RK5", "RK10", "RK20", "RK30", "RK40",
                                 "MC0","MC2", "MC4", "MC6", "MC8",
                                 "NA_internal_loops", "NA_all_loops", "NA_dens_directed",
                                 "NA_reciproc", "NA_transitivity_dir")

ss_field_sel_df <- ss_field_df %>% select("RK2","RK5", "RK10", "RK20", "RK30", "RK40",
                                 "MC0","MC2", "MC4", "MC6", "MC8", 
                                 "NA_internal_loops", "NA_all_loops", "NA_dens_directed",
                                 "NA_reciproc", "NA_transitivity_dir")

#standardize data according to priors ss_all
ss_all_nost <- ss_all_sel
ss_field_nost <- ss_field_sel_df

# for (col in 1:ncol(ss_all_sel)){
#   ss_all_sel[,col] <- (ss_all_sel[,col]-mean(ss_all_nost[,col]))/sd(ss_all_nost[,col])
#   ss_field_sel_df[,col] <- (ss_field_sel_df[,col]-mean(ss_all_nost[,col]))/sd(ss_all_nost[,col])
# }

#make field data vector
ss_field_sel <- as.numeric(ss_field_sel_df[1,])
ss_all_sel_scen <- ss_all_nost
ss_all_sel_scen$scenario <- ss_all_w_scen$scenario
```

##Cross validation for model selection
Performed to see if ABC can distinguish between the models. The confusion matrix shows on the rows the models and in the columns how they were classified. If the ABC can distinguish the different models, the values on the diagonal should be largest.  


```{r eval=FALSE, include=FALSE}
cv_bembix <- abc::cv4postpr(index=indexes, sumstat=ss_all_sel, nval=100, tols=0.0166, method='rejection')
summary(cv_bembix)
plot(cv_bembix, names.arg=c("bimodal", "BU", "CF", "CA", "PEBU", "random"), probs=TRUE)
plot(cv_bembix, names.arg=c("bimodal", "BU", "CF", "CA", "PEBU", "random"), probs=FALSE)
```

##Model selection
'postpr' function from the package abc is used to look which model is mostly selected (when summary stats lay closest to field data sumstats). 

```{r}
#100 of 60.000 -> 0.001666
# modsel_bembix <- postpr(target=ss_field, index=indexes, sumstat=ss_all,tol=0.001,method="rejection")
# summary(modsel_bembix)

modsel_bembix <- postpr(target=ss_field_sel, index=indexes, sumstat=ss_all_sel ,tol=0.01666666, method="rejection")
summary(modsel_bembix)
```

##Summary stats of models that were excepted

Here the priors (transparant red) of the summary stats are plotted and the posterios (black). The green line is the field data.

```{r, fig.height=30, fig.width=30, message=FALSE, warning=FALSE}
index_ss <- as.numeric(rownames(modsel_bembix$ss))
stats_ss <- ss_all[index_ss,]

ggplot(data=gather(stats_ss), aes(x=value)) +
  geom_histogram(aes(x=value, y=..count../sum(..count..))) + facet_wrap(~key, scales='free') +
  geom_vline(data=gather(ss_field_df[1,]), aes(xintercept=value), color='green', linetype="dashed", size=3)+
  theme(text = element_text(size=12))+
  geom_histogram(data=gather(ss_all), aes(x=value, y=..count../sum(..count..)), fill="#FF6666", alpha=0.5)
```

###Plot of the accepted summary stats with field data

Ripley's K: spatial clustering is quite well formed.  
Mark correlation: temporal clustering is not formed in the models. (Observational bias in the field or missing things in model). There seems to be a trade-off when looking at all values of RKs and MCs in the full 60.000 runs: high MCs has low RKs and vice versa.  
Network analysis are produced quite well.  
```{r, fig.height=30, fig.width=30}
#get ripley's K, Mark correlation and network analysis values for the accepted models
stats_ss_w_scen <- ss_all_sel_scen[index_ss,]
RKs <- stats_ss_w_scen %>% select("scenario","RK2", "RK5", "RK10", "RK20","RK30", "RK40")
MCs <- stats_ss_w_scen %>% select("scenario", "MC0","MC2", "MC4", "MC6", "MC8")
NAs <- stats_ss_w_scen %>% select("scenario", "NA_internal_loops", "NA_all_loops",
                                  "NA_dens_directed","NA_reciproc", "NA_transitivity_dir")

# str(RKs)
# str(MCs)
# str(NAs)

#get field summary stats
field_RK <- ss_field_nost %>%
  select("RK2","RK5", "RK10", "RK20","RK30", "RK40") %>%
  filter(row_number()==1)
field_RK$scenario <- 'field data'

field_MC <- ss_field_nost %>%
  select("MC0","MC2", "MC4", "MC6", "MC8") %>%
  filter(row_number()==1)
field_MC$scenario <- 'field data'

field_NA <- ss_field_nost %>% select("NA_internal_loops", "NA_all_loops",
                                  "NA_dens_directed","NA_reciproc", "NA_transitivity_dir") %>%
  filter(row_number()==1)

priors_RK <- ss_all_sel_scen  %>%
  select("scenario", "RK2","RK5", "RK10", "RK20", "RK30", "RK40")
priors_MC <- ss_all_sel_scen %>%
  select("scenario","MC0","MC2", "MC4", "MC6", "MC8")
priors_NA <- ss_all_sel_scen %>% select("scenario","NA_internal_loops", "NA_all_loops",
                                  "NA_dens_directed","NA_reciproc", "NA_transitivity_dir")

#change colnames to easily plot the functions correctly
colnames(RKs) <- c("scenario",'2', '5', '10', '20', '30', '40')
colnames(field_RK) <- c('2', '5', '10', '20', '30', '40',"scenario")
colnames(priors_RK) <- c('scenario','2', '5', '10', '20','30', '40')

colnames(MCs) <- c("scenario",'0', '2', '4', '6', '8')
colnames(field_MC) <- c('0','2', '4', '6', '8',"scenario")
colnames(priors_MC) <- c('scenario','0', '2', '4', '6', '8')

colnames(NAs) <- c("scenario",'internal loops','all loops','density',
                   'reciprocity', 'transitivity')
colnames(field_NA) <- c('internal loops','all loops','density',
                   'reciprocity', 'transitivity')
colnames(priors_NA) <- c("scenario",'internal loops','all loops','density',
                   'reciprocity', 'transitivity')

#give the scenario's a name again (instead of a number)  
RKs$scenario <- modsel_bembix$values
MCs$scenario <- modsel_bembix$values
NAs$scenario <- modsel_bembix$values

#Ripley's K graphs
# RKs_tidy <- RKs %>%
#   gather(distance, value, '2':'40') %>%
#   mutate(distance = as.factor(as.numeric(distance))) %>%
#   group_by(scenario, distance)
# RKs_summary <- RKs_tidy %>% summarise(mean_RK = mean(value), sd_RK = sd(value))
# 
field_RK_tidy <- field_RK %>% gather(distance, value, '2':'40') %>%
  mutate(distance = as.factor(as.numeric(distance)))
RKs_tidy <- RKs %>% gather("2","5", "10", "20", "30", "40",
                           key="distance", value="RK_value") %>%
  mutate(distance = as.factor(as.numeric(distance)))
priors_RK_tidy <- priors_RK  %>%
  gather("2", "5", "10", "20","30", "40", key="distance", value="RK_value") %>%
  mutate(distance = as.factor(as.numeric(distance)))
  # filter(scenario=="random")
random_RK <- priors_RK_tidy %>% filter(scenario=="random")
  
RK_plot <- ggplot(RKs_tidy, aes(distance, RK_value)) +
  geom_violin(fill="darkgreen",alpha=0.8, position=position_nudge(x=0.1), scale="width", width=0.5) +
  #geom_line(colour="grey", linetype="dashed", size=1.5, alpha=0.5) +
  #geom_errorbar(aes(ymin=mean_RK-sd_RK, ymax=mean_RK+sd_RK, width=0.1),
  #              position=position_dodge(width=0.5)) +
  ggtitle("Ripley's K")+
  ylab("K")+
  theme(text = element_text(size=12))+
  geom_violin(data=priors_RK_tidy, aes(x=distance, y=RK_value),
              fill="grey", position=position_nudge(x=-0.1), alpha=0.8, scale="width", width=0.5)+
  geom_violin(data=random_RK, aes(x=distance, y=RK_value),
              fill="gold", alpha=0.8, colour="gold", scale="width", width=0.5)+
  geom_point(data=field_RK_tidy, aes(x=distance, y=value), fill="royalblue1", shape=21, size=3)+
  theme_bw()
  #geom_line(data=field_RK_tidy, aes(x=distance, y=value), colour='gold', size=1.5)
RK_plot

#mark correlation function
# MCs_tidy <- MCs %>%
#   gather(distance, value_d, '2':'10') %>%
#   mutate(distance = as.factor(as.numeric(distance))) %>%
#   group_by(scenario, distance)
# MCs_summary <- MCs_tidy %>%summarise(mean_MC = mean(value_d), sd_MC = sd(value_d))
# field_MC_tidy <- field_MC %>% gather(distance, value, '2':'10')

field_MC_tidy <- field_MC %>% gather(distance, value, '0':'8') %>%
  mutate(distance = as.factor(as.numeric(distance)))
MCs_tidy <- MCs %>% gather("0","2", "4", "6", "8", key="distance", value="MC_value") %>%
  mutate(distance = as.factor(as.numeric(distance)))
priors_MC_tidy <- priors_MC  %>%
  gather("0","2", "4", "6", "8", key="distance", value="MC_value") %>%
  mutate(distance = as.factor(as.numeric(distance))) #%>%
  # filter(scenario=="random")
random_MC <- priors_MC_tidy %>% filter(scenario=="random")

MC_plot <- ggplot(MCs_tidy, aes(x=distance, y=MC_value)) +
  geom_violin(fill="darkgreen", alpha=0.8, position=position_nudge(x=0.1), scale="width", width=0.5)+
  ggtitle("Mark correlation")+
  ylab("k")+
  theme(text = element_text(size=12))+
  geom_violin(data=priors_MC_tidy, aes(x=distance, y=MC_value),
              fill="grey", position=position_nudge(x=-0.1), alpha=0.8, scale="width", width=0.5)+
  geom_violin(data=random_MC, aes(x=distance, y=MC_value),
              fill="gold", alpha=0.2, colour="gold", scale="width", width=0.5)+
  geom_point(data=field_MC_tidy, aes(x=distance, y=value), fill="royalblue1", shape=21, size=3)+
  theme_bw()

#network analysis
NAs_tidy <- NAs %>%
  gather(NA_stat, value, "internal loops":"transitivity") %>%
  group_by(scenario)

field_NA_tidy <- field_NA %>% gather(key="type_NA", value="NA_value",
                                     "internal loops":"transitivity")%>%
  group_by(type_NA)
NAs_tidy <- NAs %>% gather("internal loops":"transitivity", key="type_NA", value="NA_value")
priors_NA_tidy <- priors_NA  %>%
  gather("internal loops":"transitivity", key="type_NA", value="NA_value")
random_NA <- priors_NA_tidy %>% filter(scenario=="random")

NA_plot <- ggplot(NAs_tidy, aes(x=type_NA,y=NA_value))+
  geom_violin(fill="darkgreen", alpha=0.8, position=position_nudge(x=0.1)) +
  facet_wrap(~type_NA, scales='free', nrow=1)+
  geom_violin(data=priors_NA_tidy, aes(x=type_NA, y=NA_value),
              fill="grey", position=position_nudge(-0.1), alpha=0.8)+
  geom_hline(data=field_NA_tidy, aes(yintercept=NA_value), color='royalblue1', linetype="dashed", size=1)+
   geom_violin(data=random_NA, aes(x=type_NA, y=NA_value),
              fill="gold", alpha=0.2, colour="gold")+
  ylab("value")+
  ggtitle("Network metrics")+
  theme_bw()+
  theme(axis.text.x=element_blank(), axis.ticks.x=element_blank(),
        text = element_text(size=12), axis.title.x=element_blank())
NA_plot
  
grid.arrange(RK_plot, MC_plot, NA_plot, ncol=2, nrow=2, layout_matrix=rbind(c(1,2),c(3,3)), heights=c(1,0.5))

```


##Plot of priors and posterios parameters
Priors are in pink, posteriors in black.  

```{r, fig.height=10, fig.width=10, message=FALSE, warning=FALSE}
params_bimodal$pf <- NULL

abc_bimodal <- abc(target=ss_field, param=params_bimodal, sumstat=ss_all[ss_all_w_scen$scenario == "bimodal",] ,tol=0.01, method='rejection')

summary(abc_bimodal)

ggplot(data=gather(as.data.frame(abc_bimodal$unadj.values)), aes(x=value)) +
  geom_histogram(aes(y=..count../sum(..count..))) + facet_wrap(~key, scales='free')+
  geom_histogram(data=gather(params_bimodal), aes(x=value, y=..count../sum(..count..)), fill="#FF6666", alpha=0.5)
```