---
title: "ABC_Bembix_simple"
author: "Femke Batsleer"
date: "5 februari 2018"
output: html_document
---

#Introduction
This document analyses the Bembix model with ABC (Approximate Bayesion Computation) for IBM (Individual Based Modelling). For this, a model was built with Python, including **6 model scenario's**, ran on the UGent HPC:  
1. random  
2. BU: bottom-up  
3. PEBU: bottom-up with previous experience (staying close to previous nest)  
4. CA: conspecific attraction (attracted by digging individuals, social cue)  
5. bimodal: bimodal choice (intra-individual variation in strategy: individuals choose either to stay close to previous nest or go further with conspecific attraction for each nest)  
6. CF: choosers/followers (interindividual variation in behavioural strategy: individuals are searchers for good nesting place or followers)  
   
These runs created each 3 files: one with model parameter values, one with distances between two nests of an individual, one with the wasp, nest, coordinates and day of nestbuilding. In R (also ran on HPC), parameters were put in the correct format and **summary statistics were calculated** for each model (Ripley's K and Network analysis).  
The calculations were redone (only for real nests) for the field data and put in the same format as the summary statistics.  

This document shows ABC analysis for the Bembix IBM (with simplified amount of parameters in the models), using the package abc.  

**10.000 runs were ran for each scenario**  

#ABC analysis
##Loading data
Data are loaded, created by other R-scripts. Field data and for each scenario: parameters and summary statistics.  
The data are adjusted to make sure they have the right format for the analysis (True and False become 0 and 1, response functions get numbers etc. because everything needs to be numeric).  

###Summary statistics  
All summary stats are loaded and a restricted dataset is created to include the most important ones (the ones that we want to select models on).  
```{r, message=FALSE, warning=FALSE}
setwd("D:/fbatslee/OneDrive - UGent/WP 2 Modelling Thesis/2 - Analysis/ABC4IBM/ABC")

library(abc)
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)
library(gridExtra)

#loading the data: summary stats, params scenarios, field data summary stats
##summary stats
getwd()
ss_random <- read.table("../Data outputs/Simple 12-06-2020 INLA/Summary_stats_random_2020-08-11.txt", header=T)
ss_bimodal <- read.table("../Data outputs/Simple 12-06-2020 INLA/Summary_stats_bimodal_2020-08-11.txt", header=T)
ss_BU <- read.table("../Data outputs/Simple 12-06-2020 INLA/Summary_stats_bottom-up_2020-08-11.txt", header=T)
ss_CF <- read.table("../Data outputs/Simple 12-06-2020 INLA/Summary_stats_CF_2020-08-11.txt", header=T)
ss_CA <- read.table("../Data outputs/Simple 12-06-2020 INLA/Summary_stats_conspec-attraction_2020-08-11.txt", header=T)
ss_PEBU <- read.table("../Data outputs/Simple 12-06-2020 INLA/Summary_stats_prev-exp-BU_2020-08-11.txt", header=T)

##combine the datasets
ss_all <- plyr::rbind.fill(ss_random, ss_bimodal, ss_CA, ss_CF, ss_BU, ss_PEBU)

colnames(ss_all) <- c("scenario", "file title", 'RK2', 'RK5', 'RK10', 'RK15', 'RK20', 'RK30', 'RK40',
                      "NA_internal_loops", "NA_all_loops","NA_dens_undirected", "NA_dens_directed",
                      "NA_reciproc","NA_transitivity_und","NA_transitivity_dir")

str(ss_all)
```


###Field data
Master thesis data  
```{r}
##field data summary stats
ss_field_df <- read.table("../Field Summary stats/Summary_stats_field.txt", header=T)
ss_field_df$X0 <- NULL
ss_field_df$scenario <- NULL
ss_field_df[] <- lapply(ss_field_df[], as.numeric)
colnames(ss_field_df) <- c("mean_RK", "sd_RK", 'RK2', 'RK5', 'RK10', 'RK15', 'RK20', 'RK30', 'RK40',
                           "NA_internal_loops", "NA_all_loops","NA_dens_undirected", "NA_dens_directed",
                           "NA_reciproc","NA_transitivity_und","NA_transitivity_dir")
ss_field_df$mean_RK <- NULL
ss_field_df$sd_RK <- NULL

#make a 1D vector of it
ss_field <- as.numeric(ss_field_df[1,])

# #same for the restricted summary stats [c(1, 4:10, 13:23, 24:ncol(ss_all))]
# ss_field_res_df <- ss_field_df[c(1:3, 5:12, 19:20, 22:23, 25)]
# ss_field_res <- as.numeric(ss_field_res_df[1,])
# ss_field_res_df
```

###Parameters
Model parameters for all the runs.  
```{r, message=FALSE, warning=FALSE}
##priors/parameters of the models
#2 --> with weight day and max dens
params_random <- read.table("../Data outputs/Simple 12-06-2020 INLA/Params_random_2020-06-11.txt", header=T)
params_bimodal <- read.table("../Data outputs/Simple 12-06-2020 INLA/Params_bimodal_2020-06-11.txt", header=T)
params_BU <- read.table("../Data outputs/Simple 12-06-2020 INLA/Params_bottom-up_2020-06-11.txt", header=T)
params_CF <- read.table("../Data outputs/Simple 12-06-2020 INLA/Params_CF_2020-06-11.txt", header=T)
params_CA <- read.table("../Data outputs/Simple 12-06-2020 INLA/Params_conspec-attraction_2020-06-11.txt", header=T)
params_PEBU <- read.table("../Data outputs/Simple 12-06-2020 INLA/Params_prev-exp-BU_2020-06-11.txt", header=T)

##these dataframes are converted to numeric ones (True-->1, False-->0,...)
change_df_num <- function(df){
  df$scenario <- NULL
  df[] <- lapply(df[], as.character)
  df[df=='True'] <- '1'
  df[df=='False'] <- '0'
  df[df=='response_par'] <- '1'
  df[df=='response_type2'] <- '2'
  df[] <- lapply(df[], as.numeric)
  return(df)
}

params_bimodal <- change_df_num(params_bimodal)
params_BU <- change_df_num(params_BU)
params_CF <- change_df_num(params_CF)
params_CA <- change_df_num(params_CA)
params_PEBU <- change_df_num(params_PEBU)

##combine them in 1 dataset
params_all <- plyr::rbind.fill(params_random,params_bimodal, params_CA, params_CF, params_BU, params_PEBU)
head(params_all)
```

###Indexes
A vector has to be created with the names in of the models corresponding to the different runs.  

```{r}
#indexes -> which row corresponds with which model
indexes <- as.vector(unlist(ss_all[,1]))
str(indexes)

#keep summary stats with scenario attached to df
ss_all_w_scen <- ss_all
# ss_res_w_scen <- ss_res
#make sure the summary stats are numeric values (scenario will be replaced with 1..6)
ss_all[] <- lapply(ss_all[], as.numeric) #make sure all are numeric values
# ss_res[] <- lapply(ss_res[], as.numeric)

ss_all$scenario <- NULL #take scenario out of summary stats, to make calculations easier
ss_all$`file title` <- NULL
# ss_res$scenario <- NULL
```


####Correlation between summary stats
Within the Ripley's K, there is high correlation. But less for the high values. We should select distances that are further apart from each other. F.i. RK2, RK10, RK20, RK30, RK40.  
Within Network metrics, there is some correlation. Especially for the same metric for undirected and directed networks. These will be dropped.  
For the network metrics and Ripley's K, there is some structure. But this is expected with such high level network metrics. Variation seems to be high enough.  

```{r}
#Within Ripley's K
pairs(ss_all[sample(nrow(ss_all),1000),c(1:7)])
#within network metrics
pairs(ss_all[sample(nrow(ss_all),1000), c(8:ncol(ss_all))])

#make selection
ss_all_sel <- ss_all %>% select("RK2","RK5", "RK10", "RK20", "RK30", "RK40",
                                 "NA_internal_loops", "NA_all_loops", "NA_dens_directed",
                                 "NA_reciproc", "NA_transitivity_dir")

ss_field_sel_df <- ss_field_df %>% select("RK2","RK5", "RK10", "RK20", "RK30", "RK40",
                                 "NA_internal_loops", "NA_all_loops", "NA_dens_directed",
                                 "NA_reciproc", "NA_transitivity_dir")

#standardize data according to priors ss_all
ss_all_nost <- ss_all_sel
ss_field_nost <- ss_field_sel_df

# for (col in 1:ncol(ss_all_sel)){
#   ss_all_sel[,col] <- (ss_all_sel[,col]-mean(ss_all_nost[,col]))/sd(ss_all_nost[,col])
#   ss_field_sel_df[,col] <- (ss_field_sel_df[,col]-mean(ss_all_nost[,col]))/sd(ss_all_nost[,col])
# }

#make field data vector
ss_field_sel <- as.numeric(ss_field_sel_df[1,])
ss_all_sel_scen <- ss_all_nost
ss_all_sel_scen$scenario <- ss_all_w_scen$scenario
```

##Cross validation for model selection
Performed to see if ABC can distinguish between the models. The confusion matrix shows on the rows the models and in the columns how they were classified. If the ABC can distinguish the different models, the values on the diagonal should be largest.  

* BU, PEBU and random are almost always correctly classified.  
* CF cannot be distinguished from BU (CF uses partially the BU search mode, so not surprising)  
* Bimodal is 29.5 percent of the times correctly classified, often as CA  
* CA is most of the time (30%) classified correctly, but also often as other models, like bimodal. This is probably because the other models partially implement  

```{r eval=FALSE, include=FALSE}
cv_bembix <- abc::cv4postpr(index=indexes, sumstat=ss_all_sel, nval=100, tols=0.016666, method='rejection')
summary(cv_bembix)
plot(cv_bembix, names.arg=c("bimodal", "BU", "CF", "CA", "PEBU", "random"), probs=TRUE)
#plot(cv_bembix, names.arg=c("bimodal", "BU", "CF", "CA", "PEBU", "random"), probs=FALSE)
```

##Model selection
'postpr' function from the package abc is used to look which model is mostly selected (when summary stats lay closest to field data sumstats). 

```{r}
#100 of 60.000 -> 0.001666
# modsel_bembix <- postpr(target=ss_field, index=indexes, sumstat=ss_all,tol=0.001,method="rejection")
# summary(modsel_bembix)

modsel_bembix <- postpr(target=ss_field_sel, index=indexes, sumstat=ss_all_sel ,tol=0.016666, method="rejection")
summary(modsel_bembix)

#source("D:/fbatslee/OneDrive - UGent/Conferences/BES/ABCWorkshop/Code/R/ModelSelection.R")
```

##Summary stats of models that were excepted

Here the priors (transparant red) of the summary stats are plotted and the posterios (black). The green line is the field data.

```{r, fig.height=30, fig.width=30, message=FALSE, warning=FALSE}
index_ss <- as.numeric(rownames(modsel_bembix$ss))
stats_ss <- ss_all_sel[index_ss,]

ggplot(data=gather(stats_ss), aes(x=value)) +
  geom_histogram(aes(x=value, y=..count../sum(..count..))) + facet_wrap(~key, scales='free') +
  geom_vline(data=gather(ss_field_df[1,]), aes(xintercept=value), color='green', linetype="dashed", size=3)+
  theme(text = element_text(size=12))+
  geom_histogram(data=gather(ss_all), aes(x=value, y=..count../sum(..count..)), fill="#FF6666", alpha=0.5)
```

###Plot of the accepted summary stats with field data

Ripley's K: spatial clustering is quite well formed.  
Network analysis are produced quite well.  
```{r, fig.height=30, fig.width=30}
#get ripley's K and network analysis values for the accepted models
stats_ss_w_scen <- ss_all_sel_scen[index_ss,]
RKs <- stats_ss_w_scen %>% select("scenario","RK2", "RK5", "RK10", "RK20","RK30", "RK40")
NAs <- stats_ss_w_scen %>% select("scenario", "NA_internal_loops", "NA_all_loops",
                                  "NA_dens_directed","NA_reciproc", "NA_transitivity_dir")

# str(RKs)
# str(NAs)

#get field summary stats
field_RK <- ss_field_nost %>%
  select("RK2","RK5", "RK10", "RK20","RK30", "RK40") %>%
  filter(row_number()==1)
field_RK$scenario <- 'field data'

field_NA <- ss_field_nost %>% select("NA_internal_loops", "NA_all_loops",
                                  "NA_dens_directed","NA_reciproc", "NA_transitivity_dir") %>%
  filter(row_number()==1)

priors_RK <- ss_all_sel_scen  %>%
  select("scenario", "RK2","RK5", "RK10", "RK20", "RK30", "RK40")
priors_NA <- ss_all_sel_scen %>% select("scenario","NA_internal_loops", "NA_all_loops",
                                  "NA_dens_directed","NA_reciproc", "NA_transitivity_dir")

#change colnames to easily plot the functions correctly
colnames(RKs) <- c("scenario",'2', '5', '10', '20', '30', '40')
colnames(field_RK) <- c('2', '5', '10', '20', '30', '40',"scenario")
colnames(priors_RK) <- c('scenario','2', '5', '10', '20','30', '40')


colnames(NAs) <- c("scenario",'internal loops','all loops','density',
                   'reciprocity', 'transitivity')
colnames(field_NA) <- c('internal loops','all loops','density',
                   'reciprocity', 'transitivity')
colnames(priors_NA) <- c("scenario",'internal loops','all loops','density',
                   'reciprocity', 'transitivity')

#give the scenario's a name again (instead of a number)  
RKs$scenario <- modsel_bembix$values
NAs$scenario <- modsel_bembix$values

#Ripley's K graphs
# RKs_tidy <- RKs %>%
#   gather(distance, value, '2':'40') %>%
#   mutate(distance = as.factor(as.numeric(distance))) %>%
#   group_by(scenario, distance)
# RKs_summary <- RKs_tidy %>% summarise(mean_RK = mean(value), sd_RK = sd(value))
# 
field_RK_tidy <- field_RK %>% gather(distance, value, '2':'40') %>%
  mutate(distance = as.factor(as.numeric(distance)))
RKs_tidy <- RKs %>% gather("2","5", "10", "20", "30", "40",
                           key="distance", value="RK_value") %>%
  mutate(distance = as.factor(as.numeric(distance)))
priors_RK_tidy <- priors_RK  %>%
  gather("2", "5", "10", "20","30", "40", key="distance", value="RK_value") %>%
  mutate(distance = as.factor(as.numeric(distance)))
  # filter(scenario=="random")
random_RK <- priors_RK_tidy %>% filter(scenario=="random")
  
RK_plot <- ggplot(RKs_tidy, aes(distance, RK_value)) +
  geom_violin(fill="darkgreen", color="darkgreen", alpha=0.8, position=position_nudge(x=0.1), scale="width", width=0.5) +
  #geom_line(colour="grey", linetype="dashed", size=1.5, alpha=0.5) +
  #geom_errorbar(aes(ymin=mean_RK-sd_RK, ymax=mean_RK+sd_RK, width=0.1),
  #              position=position_dodge(width=0.5)) +
  ggtitle("Ripley's K")+
  ylab("K")+
  theme(text = element_text(size=12))+
  geom_violin(data=priors_RK_tidy, aes(x=distance, y=RK_value),
              fill="grey", color="grey", position=position_nudge(x=-0.1), alpha=0.8, scale="width", width=0.5)+
  geom_violin(data=random_RK, aes(x=distance, y=RK_value),
              fill="gold", alpha=0.8, colour="gold", scale="width", width=0.5)+
  geom_point(data=field_RK_tidy, aes(x=distance, y=value), fill="royalblue1", shape=21, size=3)+
  theme_bw()
  #geom_line(data=field_RK_tidy, aes(x=distance, y=value), colour='gold', size=1.5)
RK_plot


#network analysis
NAs_tidy <- NAs %>%
  gather(NA_stat, value, "internal loops":"transitivity") %>%
  group_by(scenario)

field_NA_tidy <- field_NA %>% gather(key="type_NA", value="NA_value",
                                     "internal loops":"transitivity")%>%
  group_by(type_NA)
NAs_tidy <- NAs %>% gather("internal loops":"transitivity", key="type_NA", value="NA_value")
priors_NA_tidy <- priors_NA  %>%
  gather("internal loops":"transitivity", key="type_NA", value="NA_value")
random_NA <- priors_NA_tidy %>% filter(scenario=="random")

NA_plot <- ggplot(NAs_tidy, aes(x=type_NA,y=NA_value))+
  geom_violin(fill="darkgreen", color="darkgreen", alpha=0.8, position=position_nudge(x=0.1)) +
  facet_wrap(~type_NA, scales='free', nrow=1)+
  geom_violin(data=priors_NA_tidy, aes(x=type_NA, y=NA_value),
              fill="grey", color="grey", position=position_nudge(-0.1), alpha=0.8)+
  geom_hline(data=field_NA_tidy, aes(yintercept=NA_value), color='royalblue1', linetype="dashed", size=1)+
   geom_violin(data=random_NA, aes(x=type_NA, y=NA_value),
              fill="gold", alpha=0.2, colour="gold")+
  ylab("value")+
  ggtitle("Network metrics")+
  theme_bw()+
  theme(axis.text.x=element_blank(), axis.ticks.x=element_blank(),
        text = element_text(size=12), axis.title.x=element_blank())
NA_plot
  
grid.arrange(RK_plot, NA_plot, ncol=1, nrow=2, layout_matrix=rbind(c(1,1),c(2,2)), heights=c(1,0.75))

```


##Plot of priors and posterios parameters
Priors are in pink, posteriors in black.  

* active: most of the time, the parameter active (if the individuals should take into account only active digging wasps or not) was mostly taken false: they take into account all nests most often.  
* bi_CA: when they choose to stay close to their previous nest, they are not searching with the social cue search mode, but most often with the bottom-up search mode (bi_CA indicated if they, staying close to previous nest, would search with social cue or with bottom-up).  
* CA_day: was mostly taken true, so they weighted the surrounding nest by the day they were made when using the social cue.  
*  CA_w_env: the environment was most often taken into account as well: so they searched both with the social cue, weighted by their own assesment of the environment.  
* response function: was mostly parabolic (0 parabolic, 1 type-2 fuction): real density dependent nest choice behaviour.  

***
>Overall, posterior parameters look as expected.

```{r, fig.height=10, fig.width=10, message=FALSE, warning=FALSE}
params_bimodal$pf <- NULL

abc_bimodal <- abc(target=ss_field, param=params_bimodal, sumstat=ss_all[ss_all_w_scen$scenario == "bimodal",] ,tol=0.01, method='rejection')

summary(abc_bimodal)

ggplot(data=gather(as.data.frame(abc_bimodal$unadj.values)), aes(x=value)) +
  geom_histogram(aes(y=..count../sum(..count..))) + facet_wrap(~key, scales='free')+
  geom_histogram(data=gather(params_bimodal), aes(x=value, y=..count../sum(..count..)), fill="#FF6666", alpha=0.5)
```