---
title: "ABC_detailed_model_selection"
author: "Femke Batsleer"
date: "17 december 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#ABC analysis
two methods are used, with abc package and with flexible script from Sibly's group. Only real difference applicable here is that during model selection different scaling is used (abc package: mad, median absolute deviation; script: sd)
```{r, message=FALSE, warning=FALSE}
setwd("D:/fbatslee/OneDrive - UGent/WP 2 Modelling Thesis/2 - Analysis/ABC4IBM/ABC")

library(abc)
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)
library(gridExtra)

#loading the data: summary stats, params scenarios, field data summary stats
##summary stats
getwd()
ss_random <- read.table("../Data outputs/Simple 12-06-2020 INLA/Summary_stats_random_2020-08-11.txt", header=T)
ss_bimodal <- read.table("../Data outputs/Simple 12-06-2020 INLA/Summary_stats_bimodal_2020-08-11.txt", header=T)
ss_BU <- read.table("../Data outputs/Simple 12-06-2020 INLA/Summary_stats_bottom-up_2020-08-11.txt", header=T)
ss_CF <- read.table("../Data outputs/Simple 12-06-2020 INLA/Summary_stats_CF_2020-08-11.txt", header=T)
ss_CA <- read.table("../Data outputs/Simple 12-06-2020 INLA/Summary_stats_conspec-attraction_2020-08-11.txt", header=T)
ss_PEBU <- read.table("../Data outputs/Simple 12-06-2020 INLA/Summary_stats_prev-exp-BU_2020-08-11.txt", header=T)

##combine the datasets
ss_all <- plyr::rbind.fill(ss_random, ss_bimodal, ss_CA, ss_CF, ss_BU, ss_PEBU)

colnames(ss_all) <- c("scenario", "file title", 'RK2', 'RK5', 'RK10', 'RK15', 'RK20', 'RK30', 'RK40',
                      "NA_internal_loops", "NA_all_loops","NA_dens_undirected", "NA_dens_directed",
                      "NA_reciproc","NA_transitivity_und","NA_transitivity_dir")

str(ss_all)
```


###Field data
Master thesis data  
```{r}
##field data summary stats
ss_field_df <- read.table("../Field Summary stats/Summary_stats_field.txt", header=T)
ss_field_df$X0 <- NULL
ss_field_df$scenario <- NULL
ss_field_df[] <- lapply(ss_field_df[], as.numeric)
colnames(ss_field_df) <- c("mean_RK", "sd_RK", 'RK2', 'RK5', 'RK10', 'RK15', 'RK20', 'RK30', 'RK40',
                           "NA_internal_loops", "NA_all_loops","NA_dens_undirected", "NA_dens_directed",
                           "NA_reciproc","NA_transitivity_und","NA_transitivity_dir")

ss_field_df$mean_RK <- NULL
ss_field_df$sd_RK <- NULL
#make a 1D vector of it
ss_field <- as.numeric(ss_field_df[1,])

# #same for the restricted summary stats [c(1, 4:10, 13:23, 24:ncol(ss_all))]
# ss_field_res_df <- ss_field_df[c(1:3, 5:12, 19:20, 22:23, 25)]
# ss_field_res <- as.numeric(ss_field_res_df[1,])
# ss_field_res_df
```

###Parameters
Model parameters for all the runs.  
```{r, message=FALSE, warning=FALSE}
##priors/parameters of the models
#2 --> with weight day and max dens
params_random <- read.table("../Data outputs/Simple 12-06-2020 INLA/Params_random_2020-06-11.txt", header=T)
params_bimodal <- read.table("../Data outputs/Simple 12-06-2020 INLA/Params_bimodal_2020-06-11.txt", header=T)
params_BU <- read.table("../Data outputs/Simple 12-06-2020 INLA/Params_bottom-up_2020-06-11.txt", header=T)
params_CF <- read.table("../Data outputs/Simple 12-06-2020 INLA/Params_CF_2020-06-11.txt", header=T)
params_CA <- read.table("../Data outputs/Simple 12-06-2020 INLA/Params_conspec-attraction_2020-06-11.txt", header=T)
params_PEBU <- read.table("../Data outputs/Simple 12-06-2020 INLA/Params_prev-exp-BU_2020-06-11.txt", header=T)

##these dataframes are converted to numeric ones (True-->1, False-->0,...)
change_df_num <- function(df){
  scenario <- df$scenario
  df$scenario <- NULL
  df[] <- lapply(df[], as.character)
  df[df=='True'] <- '1'
  df[df=='False'] <- '0'
  df[df=='response_par'] <- '1'
  df[df=='response_type2'] <- '2'
  df[] <- lapply(df[], as.numeric)
  df$scenario <- scenario
  return(df)
}

params_bimodal <- change_df_num(params_bimodal)
params_BU <- change_df_num(params_BU)
params_CF <- change_df_num(params_CF)
params_CA <- change_df_num(params_CA)
params_PEBU <- change_df_num(params_PEBU)

##combine them in 1 dataset
params_all <- plyr::rbind.fill(params_random,params_bimodal, params_CA, params_CF, params_BU, params_PEBU)
head(params_all)
```

###Indexes
A vector has to be created with the names in of the models corresponding to the different runs.  

```{r}
#indexes -> which row corresponds with which model
indexes <- as.vector(unlist(ss_all[,1]))
str(indexes)

#keep summary stats with scenario attached to df
ss_all_w_scen <- ss_all
# ss_res_w_scen <- ss_res
#make sure the summary stats are numeric values (scenario will be replaced with 1..6)
ss_all[] <- lapply(ss_all[], as.numeric) #make sure all are numeric values
# ss_res[] <- lapply(ss_res[], as.numeric)

ss_all$scenario <- NULL #take scenario out of summary stats, to make calculations easier
ss_all$`file title` <- NULL
# ss_res$scenario <- NULL
```


####Selection of summary stats

```{r}
#make selection
ss_all_sel <- ss_all %>% select("RK2","RK5", "RK10", "RK20", "RK30", "RK40",
                                 "NA_internal_loops", "NA_all_loops", "NA_dens_directed",
                                 "NA_reciproc", "NA_transitivity_dir")

ss_field_sel_df <- ss_field_df %>% select("RK2","RK5", "RK10", "RK20", "RK30", "RK40",
                                 "NA_internal_loops", "NA_all_loops", "NA_dens_directed",
                                 "NA_reciproc", "NA_transitivity_dir")

#standardize data according to priors ss_all
ss_all_nost <- ss_all_sel
ss_field_nost <- ss_field_sel_df

# for (col in 1:ncol(ss_all_sel)){
#   ss_all_sel[,col] <- (ss_all_sel[,col]-mean(ss_all_nost[,col]))/sd(ss_all_nost[,col])
#   ss_field_sel_df[,col] <- (ss_field_sel_df[,col]-mean(ss_all_nost[,col]))/sd(ss_all_nost[,col])
# }

#make field data vector
ss_field_sel <- as.numeric(ss_field_sel_df[1,])
ss_all_sel_scen <- ss_all_nost
ss_all_sel_scen$scenario <- ss_all_w_scen$scenario
```

# Model selection and Bayes factors for 2 seperate groups of summary stats
Ripley's K: most support for bimodal; CA, CF, bimodal all three selected  
Bimodal and CA both substantially better than CF; random, bimodal, CA, CF selected.
Network analysis: decisive (only) support for bimodal!

```{r}
# #script from Richard Sibly's group (Elske van der Vaart)
# source("D:/fbatslee/OneDrive - UGent/Conferences/BES/ABCWorkshop/Code/R/ModelSelection.R")

params_all$rownumber <- rownames(params_all)
#model selection only on Ripley's K
# RK.abc <- create.abcSel(target = ss_field_sel[1:6], indexes=indexes, results=ss_all_sel[,1:6], rate=0.016666)
# summary(RK.abc)
RK.abc <- postpr(target = ss_field_sel[1:6], index=indexes, sumstat=ss_all_sel[,1:6], tol=0.016666, method="rejection")
summary(RK.abc)
# RK.m <- postpr(target = ss_field_sel[1:6], index=indexes, sumstat=ss_all_sel[,1:6], tol=0.00166, method="rejection")
# summary(RK.m)
indexRK_ss <- as.numeric(rownames(RK.abc$ss))
params_all$accepted_RK <- ifelse(params_all$rownumber %in% indexRK_ss, TRUE, FALSE)


#model selection on on network analysis
# NA.abc <- create.abcSel(target = ss_field_sel[12:16], indexes=indexes, results=ss_all_sel[,12:16], rate=0.016666)
# summary(NA.abc)
NA.abc <- postpr(target = ss_field_sel[7:11], index=indexes, sumstat=ss_all_sel[,7:11], tol=0.016666, method="rejection")
summary(NA.abc)

indexNA_ss <- as.numeric(rownames(NA.abc$ss))
params_all$accepted_NA <- ifelse(params_all$rownumber %in% indexNA_ss, TRUE, FALSE)
```


# Model selection for all summary stats together
1000 best models (of 60.000, which is 1.6%) chosen.
76.9% included environment (when CA, CF, bimodal).
35 times prev_exp BU was chosen, without CA. CA and prev-exp_Bu have Bayes factor around 1, so equally good.
As more runs got accepted that only include some kind of CA (225), than only with environment (35), the conspecific attraction seems to be the stronger mechanism. But both needed to eventually come to the 'best' pattern (100 best are all bimodal).
```{r}
# bembix.abc <- create.abcSel(target = ss_field_sel, indexes = indexes, results = ss_all_sel, rate = 0.01666)
# summary(bembix.abc)

bembix.abc <- postpr(target = ss_field_sel, index=indexes, sumstat=ss_all_sel, tol=0.016666, method="rejection")
summary(bembix.abc)

#indexes accepted
index_ss <- as.numeric(rownames(bembix.abc$ss))
params_all$accepted <- ifelse(params_all$rownumber %in% index_ss, TRUE, FALSE)

#select, selected parameters
#params_all$accepted <- bembix.abc$accepted
params_accepted <- params_all %>% filter(accepted==TRUE) %>%
  mutate_at(c("bi_CA", "active", "CA_w_env", "CA_day", "response_func"), as.numeric) %>%
  plyr::count(c("bi_CA"))
params_accepted

params_table <- params_all %>%
  group_by(scenario, CA_w_env, accepted) %>%
  summarise(count=n()) %>%
  spread(key=accepted, value=count) %>%
  mutate(accepted_perc = `TRUE`/10) %>%
  select(scenario, CA_w_env,accepted_perc)
params_table

params_table_RK <- params_all %>%
  group_by(scenario, CA_w_env, accepted_RK) %>%
  summarise(count=n()) %>%
  spread(key=accepted_RK, value=count) %>%
  mutate(accepted_RK_perc = `TRUE`/10) %>%
  select(scenario, CA_w_env, accepted_RK_perc)
params_table_RK

params_table_NA <- params_all %>%
  group_by(scenario, CA_w_env, accepted_NA) %>%
  summarise(count=n()) %>%
  spread(key=accepted_NA, value=count) %>%
  mutate(accepted_NA_perc = `TRUE`/10) %>%
  select(scenario, CA_w_env, accepted_NA_perc)
params_table_NA

#put them all in one table
table_perc <- params_table %>%
  left_join(params_table_RK, by=c("scenario", "CA_w_env")) %>%
  left_join(params_table_NA, by=c("scenario", "CA_w_env"))
table_perc

# params_accepted_pack <- params_all[indexes_accepted_pack,]%>%
#   mutate_at(c("bi_CA", "active", "CA_w_env", "CA_day", "response_func"), as.numeric) %>%
#   plyr::count(c("CA_w_env"))
# params_accepted_pack
```

#Get combination of other parameters
```{r}
table_active <- params_all %>%
  filter(accepted==TRUE) %>%
  group_by(scenario, CA_w_env, active) %>%
  summarise(count=n()) %>%
  spread(key=active, value=count)%>%
  mutate(active_perc = `1`/(`0`+`1`)) %>%
  select(scenario, CA_w_env, active_perc)
table_active

table_day <- params_all %>%
  filter(accepted==TRUE) %>%
  group_by(scenario, CA_w_env, CA_day) %>%
  summarise(count=n()) %>%
  spread(key=CA_day, value=count)%>%
  mutate(day_perc = `1`/(`0`+`1`)) %>%
  select(scenario, CA_w_env, day_perc)
table_day

table_nearCA <- params_all %>%
  filter(accepted==TRUE) %>%
  group_by(scenario, CA_w_env, bi_CA) %>%
  summarise(count=n()) %>%
  spread(key=bi_CA, value=count)%>%
  mutate(nearCA_perc = `1`/(`0`+`1`))%>%
  select(scenario, CA_w_env, nearCA_perc)
table_nearCA

table_params_perc <- table_active %>%
  left_join(table_day, by=c("scenario", "CA_w_env")) %>%
  left_join(table_nearCA, by=c("scenario", "CA_w_env"))
table_params_perc

#amount of runs of each scenario
table_params_runs <- params_all %>%
  group_by(scenario, CA_w_env) %>%
  summarise(count=n())
table_params_runs

# table_active <- params_all %>%
#   filter(accepted==TRUE) %>%
#   group_by(scenario, CA_w_env, active, CA_day) %>%
#   summarise(count=n()) %>%
#   spread(key=active, value=count)%>%
#   mutate(active_perc = `1`/(`0`+`1`)) %>%
#   select(scenario, CA_w_env, active_perc)
# View(table_active)
```


#only compare bimodal with env (both), bimodal without env (CA), PEBU (env)
About equal amount of runs taken for 3 scenario's (5000). If different ones taken for PEBU, gives slightly different outcome. But qualitatively always the same: PEBU much worse than bimodal without env, full bimodal always way best. So overall: internal mechanisms more important for pattern than bottom-up template


```{r}
#make new columns to be able to select scenarios in params bimodal and PEBU
params_contrast <- params_all %>% mutate(env_log = as.logical(CA_w_env)) %>%
  mutate(contrast_scen = as.logical(scenario=='bimodal' | scenario=='prev-exp-BU')) %>%
  mutate(new_scen = paste(scenario, CA_w_env, sep="_"))
#select scenarios in ss_all_sel
ss_contrast_sel <- ss_all_sel[params_contrast$contrast_scen,] %>% slice(1:10000,12000:17000)
indexes_contrast <- as.vector(unlist(params_contrast[params_contrast$contrast_scen,][c(1:10000,12000:17000),"new_scen"]))

model_contrast <- postpr(target=ss_field_sel, index=indexes_contrast, sumstat=ss_contrast_sel ,tol=0.00667, method="rejection")
summary(model_contrast)

table(indexes_contrast)

#only RK
model_contrast_RK <- postpr(target=ss_field_sel[1:6], index=indexes_contrast, sumstat=ss_contrast_sel[1:6] ,tol=0.00667, method="rejection")
summary(model_contrast_RK)


#only NA -> here bimodal without env seems better, internal dynamics most important
model_contrast_NA <- postpr(target=ss_field_sel[7:11], index=indexes_contrast, sumstat=ss_contrast_sel[7:11] ,tol=0.00667, method="rejection")
summary(model_contrast_NA)

cv_contrast <- abc::cv4postpr(index=indexes_contrast, sumstat=ss_contrast_sel, nval=50, tols=0.0166, method='rejection')
summary(cv_contrast)
plot(cv_contrast, names.arg=c("bimodal_0", "bimodal_1", "PEBU"), probs=TRUE)
```
